A Computational Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-03-30 paper/paper.tex
pass 1/2

\section{Known Neurobiological Properties of the Default Mode Network}
We begin by a deconstruction the overall DMN function
by the mechanistic relevance of each individual network node
based on published findings from neuroscience experiments.
%
\subsection{The posteriormedial cortex: Global monitoring and integration}
The posteriormedial cortex (PMC) and
the dorsomedial prefrontal cortex (dmPFC) within the DMN
are responsible for the highest metabolic turn-over
of glucose energy consumption in humans \citep{raichle2001pnas}.
The PMC myelinates relatively late during postnatal development in monkeys
\citep{goldman1987development}, generally considered to
be a sign of phylogenetic sophistication (Flechsig, 1920).
Physiological and disturbed metabolic fluctuations in the
human PMC have been repeatedly related to
phenomena of changed conscious awareness,
including anesthesia,
sleep, and forms of coma \citep{cavanna2006precuneus}.
%
The PMC has long been speculated to reflect constant computation of
the environmental statistics and its internal representation
as an "inner minds eye" \citep{cavanna2006precuneus}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from damage in the bilateral parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in the gathering
of information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity modalities
with the amygdala involved in significance evaluation and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the human ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.
%
Indeed,
electrophysiological recordings in animals implicated the PMC in
strategic selection \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
and outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
while its retrosplenial portion was
more specifically implicated in approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would persue an exploratory or exploitative
behavioral strategy in a complex food foraging task \citep{pearson2009neurons}.
Single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.

@@

\subsection{The prefrontal cortex: Stimulus-value assocation and action consideration}
The dmPFC subserves predominantly ambiguous amodal processes
across time, space, and content domains in
sensory-independent top-down pathways.
This DMN node has been described as a “mental sketchpad”
\citep{goldman1996prefrontal},
potentially implicated in de-novo generation and binding
of meaning representations instructed by stored semantics and memories.
%
The dmPFC may thus enable inference, representation, and assessment
of one's own and other individuals' action and thoughts.
Generally,
patients with neurological lesions in the prefrontal cortex
are known to adapt to novel situations and stimuli with difficulty
\citep{stuss1986frontal}.
Specifically, neural activity in human dmPFC
reflected expectations about other peoples actions and errors thereof.
dmPFC activity indeed explained the proficiency decline
of inferring other peoples' thoughts in the elderly \citep{moran2012social}.
Some dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
Also the topographically neighboring dorsal anterior cingulate cortex
has been linked to computing values and efforts of
persisting a behavioral plan versus switching the
environmental context in several lesion studies \citep{kolling2016value}.
%
Such highly abstract neural computations necessarily rely on the
generation of probabilistic internal information drawing from
episodic memory retrieval, generative construction processes,
and explicit knowledge of the external world.
%
According to computational neuroimaging experiments,
the dmPFC activity preferentially models stimulus-value associations of
possible actions that are not actually executed,
whereas the vmPFC activity models value in the environment that leads performed actions
\citep{nicolle2012agent}.

The ventromedial prefrontal cortex (vmPFC) subserves
less ambiguous subjective-value-related evaluative processes and
reward-informed risk estimation of self-relevant environmental stimuli.
This DMN node is more closely associated with
orchestrating adapted behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on model-based value representations
\citep{doherty2015structure}.
Quantative lesion findings across 344 human individuals confirmed
a gross impairment in value-based decision making
\citep{glascher2012lesion}.
The vmPFC is preferentially connected with limbic and reward-related areas.
The vmPFC has been observed to have monosynaptical connections
with the nucleus accumbens
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Additionally, the gray-matter volume of the vmPFC and nucleus accumbens
correlated with indices of value-guided behavior and social reward attitudes
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals from dopaminergic pathways
\citep{schultz1998predictive}
that not only modulate motivated behavior towards basic survival needs but also
subserve RL in humans more broadly \citep{doherty2015structure}.
This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys
\citep{croxson2005quantitative} that
quantified this brain regions to
be substantially more likely connected to the vmPFC than dmPFC in both species.
Two functional connectivity modalities in humans strongly connected
the vmPFC with the nucleus accumbens, hippocampus (HC),
and posteriomedial cortex.
%
The vmPFC is often proposed to be involved in (external) emotional
reactions and own (visceral) arousal.
Real or imaginged bodily states could be mapped in the vmPFC
as a bioregulatory disposition governing cognition
and decision making \citep{damasio1996somatic}
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
valuation
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in vmPFC.
Such a \textit{cognitive map} of action space was argued to encode
the current task state even when states are unobservable from sensory input,
which was shown to be critical for behavior.
Additionally,
independent whole-brain analyses from structural
neuroimaging studies related the gray-matter volume GMV of the vmPFC,
more consistently than any other
brain region, to indices of
social competence and social network complexity,
among the most complicated decision that humans and monkeys take
\citep{behrens2009computation}.
%

@@

Reinforcement learning is a problem-solving technique in which, through their interactions with an external environment,
an agent learns to reach a goal, optimize a reward signal, or minimize some cost,
in an iterative trial-by-error fashion.
At any moment, each action $a$ taken triggers a changes in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
$r = r(s, a)$ (or regret!, since it can be negative) collected by the
agent.
In this view, the environment is partially controlled by
the action of the agent and the reward should be thought
of as satisfaction --or punishment-- accompanying the execution of
an action, and may be delayed, only becoming substantial when accumulated over an extended period of time.
The environment is generally taken as stochastic, i.e unpredictable, changing in random ways. In addition,
it is only \textit{partially observed},
in the sense that only part of the current state is observed by
the agent \citep{starkweather2017dopamine}.
This unpredictability of the environment
\textbf{XXX Explain this sentence better because it is important.}
is realistic in a model which sets out
to explain DMN function. Indeed, it is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensory inputs
\citep{antrobus1966studies, pope1978regulation},
the more stimulus-independent thoughts occur, and
the higher the neural activity in the DMN.
Conversely, the more the currently executed task is unknown and unpracticed,
the lower the DMN activity
\citep{filler1973daydreaming, teasdale1995stimulus}.
%
These ``offline'' processes may however contribute to
optimizing control of the organism.
A DMN framework incorporating reinforcement learning
can naturally embed human behavior
into the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules
\citep{dayan2008decision}.
The implication of the DMN in a diversity of human cognitive processes
can be parsimoneously explained as prediction error minimization
coupled with probabilistic simulations in form of mental scenes,
thus calibrating action portfolios and
maximizing reward outcome at various time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.

\subsection{Partially Observable Markov Decision Processes}
It is our view that at a sufficiently coarse scale, the brain --and therefore the DMN in particular--
is a physical system governed by the laws of
physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in phase space.
If the process has $memory$ such that the next state depends not only on the current state but also on some finite
number of past states,
rational probabilistic planning can be reformulated
as a standard memoryless Markov process by simply expanding the
definition of the state $s$ to include elements of the past.

In artificial intelligence (AI), a popular abstract model for
multi-step decision processes in such an environment are so-called
\textit{Partially Observable Markov Decision Processes (POMDPs)}.
A POMDP models a decision process in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state. Instead, it tries to optimize a \textit{subjective} reward
signal. This is performed alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.

@@

In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of world.
These can be assumptions about the physical laws governing the agent's world or constraints on the state space and transitions between states.
For instance, if a rat in a maze knows
that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
% A core property of human intelligence has been proposed underly to improve
% on an expected utility outcomes as a strategy for action choice in uncertain
% environments \citep{gershman2015computational}.
Our rat might represent such knowledge about the world as:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}

@@

\subsection{Long-term rewards and action policies}
The behavior of the agent is governed by some kind of \textit{policy},
which maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades as follows
\begin{eqnarray*}
  \begin{split}
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect rewards }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect rewards }R_1 = r(s_1, a_1)\\
    \ldots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect rewards }R_{t} = r(s_{t}, a_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken in the present moment $t$
might have repercussions long into the future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account the future. A standard for
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the brain ? You mean where in the DMN!
  then the answer is in the vmPFC which is connected to the NAc}
  The amygdala is known to be involved in significance evaluation) and
  the nucleus accumbens (NAc) is involved in reward evaluation.
\end{mdframed}
  The  goal of RL is then to adapt this policy so as to  maximize $G^\pi$. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Taking $\gamma = 0$   corresponds to a perfectly myopic agent who is solely concerned about their
immediate rewards and has no horizon for the future, which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is thus necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Thus, the agent is more considering of rewards in to come in a short future, paying lesser attention to
rewards to come in at a later time.
% given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can then be seen as calibrating
the delay discounting behavior of the intelligent agent,
that is the decision related to reward-delay trade-off schedules.
% It is also important to appreciate that a stochastic policy estimation
% is more advantageous in many RL settings.
% For instance, a deterministic strategy in playing rock-paper-scissors can
% be easily exploited and a uniformly random choice is optimal.
% \textbf{XXX I love this example :-)}

\subsection{The Q-value: A goodness measure for actions}
The purpose of the DMN can then be
formulated as finding a policy $\pi$ for the agent which maximizes the
expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a,a_1 \sim \pi,a_2 \sim \pi,\ldots].
\end{equation}
In other words,
the Q-value $Q^{\pi}(s,a)$ corresponds to the expected reward
over all possible action trajectories in which
the agent starts off starts at with an environment in state
$s$, chooses action $a$,
and then follows the policy $\pi$ thereafter to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action; it answers the question
"What is the expected utiltiy of taking action $a$ in this situation ?".
When a choice of actions is available, the $Q^{\pi}(s,a)$
provides a formalization of optimal behavior,
potentially driven by the DMN.

\subsection{Computing optimal behavior}
By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general
due to the enormous number of possible action
trajectories the agent might need to
consider simultaneously.
% \begin{mdframed}
%   \textbf{N.B.:}
%   If the Bellman-based computational ideas have a corresponding
%   neurobiological implementation, this provides a framework of how
%   optimal behavior could be computable in the human brain.
% \end{mdframed}
One popular solution concept is called \text{Q-learning}. It remedies
this issue is to only consider the sub-family of
\textit{deterministic policies}, which map each state to the best single best action to
take at the state. Such policies define deterministic functions from states to actions.

\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992), generates a
\textit{greedy policy}
\begin{equation}
  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
  \label{eq:qlearning}
\end{equation}
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
The \textit{Bellman equation} \citep{sutton1998reinforcement} then takes the simple form
\begin{equation}
  \begin{split}
    Q^\pi(s, a) &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
    \\
    &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))],
  \end{split}
  \label{eq:bellman}
\end{equation}
a \textit{fixed-point} equation which provides a recursive decomposition of optimal behavior.
A complicated dynamic programming optimization can be
broken into simpler sub-problems at different time points,
which has often been closely related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value,
to guide action towards a better state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Notice that in \eqref{eq:bellman}, the sampling is done over only things which
\textbf{XXX Ahhmm, please make a bigger effort here}
depend on the environment, and so can be learned off-policy by observing state-transitions
triggered by another behavioral policy, which can be stochastic.

>>>>

\section{Known Neurobiological Properties of the Default Mode Network}
We begin by a deconstruction the overall DMN function
by the mechanistic relevance of each individual network node
based on published findings from neuroscience experiments.
%
\subsection{The posteromedial cortex: Global monitoring and integration}
The posteriormedial cortex (PMC) and
the dorsomedial prefrontal cortex (dmPFC) within the DMN
are responsible for the highest metabolic turn-over
of glucose energy consumption in humans \citep{raichle2001pnas}.
The PMC myelinates relatively late during postnatal development in monkeys
\citep{goldman1987development}, generally considered to
be a sign of phylogenetic sophistication (Flechsig, 1920).
Physiological and disturbed metabolic fluctuations in the
human PMC have been repeatedly related to
phenomena of changed conscious awareness,
including anesthesia,
sleep, and forms of coma \citep{cavanna2006precuneus}.
%
The PMC has long been speculated to reflect constant computation of
the environmental statistics and its internal representation
as an "inner minds eye" \citep{cavanna2006precuneus}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from damage in the bilateral parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in the gathering
of information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity modalities
with the amygdala involved in significance evaluation and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the human ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.
%
Indeed,
electrophysiological recordings in animals implicated the PMC in
strategic selection \citep{pearson2009neurons},
risk assessment \citep{mccoy2005risk},
and outcome-contingent behavioral modulation \citep{hayden2009electrophysiological},
while its retrosplenial portion was
more specifically implicated in approach-avoidance behavior
\citep{vann2009does}.
Neuron spiking activity in the PMC allowed distinguishing
whether a monkey would persue an exploratory or exploitative
behavioral strategy in a complex food foraging task \citep{pearson2009neurons}.
Single-cell recordings in the monkey PMC
demonstrated this brain region's sensitivity to
subjective target utility \citep{mccoy2005risk} and integration
across individual decision-making instances \citep{pearson2009neurons}.

@@

\subsection{The prefrontal cortex: Stimulus-value association and action consideration}
The dmPFC subserves predominantly ambiguous amodal processes
across time, space, and content domains in
sensory-independent top-down pathways.
This DMN node has been described as a “mental sketchpad”
\citep{goldman1996prefrontal},
potentially implicated in de-novo generation and binding
of meaning representations instructed by stored semantics and memories.
%
The dmPFC may thus enable inference, representation, and assessment
of one's own and other individuals' action and thoughts.
Generally,
patients with neurological lesions in the prefrontal cortex
are known to adapt to novel situations and stimuli with difficulty
\citep{stuss1986frontal}.
Specifically, neural activity in human dmPFC
reflected expectations about other peoples actions and errors thereof.
dmPFC activity indeed explained the proficiency decline
of inferring other peoples' thoughts in the elderly \citep{moran2012social}.
Some dmPFC neurons in macaque monkeys exhibited a preference
for processing others', rather than own, behavior
with fine-grained adjustment of contextual circumstances \citep{yoshida2010neural}.
Also the topographically neighboring dorsal anterior cingulate cortex
has been linked to computing values and efforts of
persisting a behavioral plan versus switching the
environmental context in several lesion studies \citep{kolling2016value}.
%
Such highly abstract neural computations necessarily rely on the
generation of probabilistic internal information drawing from
episodic memory retrieval, generative construction processes,
and explicit knowledge of the external world.
%
According to computational neuroimaging experiments,
the dmPFC activity preferentially models stimulus-value associations of
possible actions that are not actually executed,
whereas the vmPFC activity models value in the environment that leads performed actions
\citep{nicolle2012agent}.

The ventromedial prefrontal cortex (vmPFC) subserves
less ambiguous subjective-value-related evaluative processes and
reward-informed risk estimation of self-relevant environmental stimuli.
This DMN node is more closely associated with
orchestrating adapted behavior by bottom-up-driven
processing of “what matters now”,
probably drawing on model-based value representations
\citep{doherty2015structure}.
Quantative lesion findings across 344 human individuals confirmed
a gross impairment in value-based decision making
\citep{glascher2012lesion}.
The vmPFC is preferentially connected with limbic and reward-related areas.
The vmPFC has been observed to have monosynaptical connections
with the nucleus accumbens
in axonal tracing studies in monkeys \citep{haber1995orbital}.
Additionally, the gray-matter volume of the vmPFC and nucleus accumbens
correlated with indices of value-guided behavior and social reward attitudes
\citep{lebreton2009automatic}.
NAc activity is thought to reflect reward prediction signals from dopaminergic pathways
\citep{schultz1998predictive}
that not only modulate motivated behavior towards basic survival needs but also
subserve RL in humans more broadly \citep{doherty2015structure}.
This is consistent with diffusion MRI (dMRI) tractography in humans and monkeys
\citep{croxson2005quantitative} that
quantified this brain regions to
be substantially more likely connected to the vmPFC than dmPFC in both species.
Two functional connectivity modalities in humans strongly connected
the vmPFC with the nucleus accumbens, hippocampus (HC),
and posteromedial cortex.
%
The vmPFC is often proposed to be involved in (external) emotional
reactions and own (visceral) arousal.
Real or imaginged bodily states could be mapped in the vmPFC
as a bioregulatory disposition governing cognition
and decision making \citep{damasio1996somatic}
In neuroeconomic studies of human decision making,
the vmPFC consistently reflects an individual’s subjective
valuation
\citep{behrens2008associative}.
This may be why performance within and across participants
was related to state encoding in vmPFC.
Such a \textit{cognitive map} of action space was argued to encode
the current task state even when states are unobservable from sensory input,
which was shown to be critical for behavior.
Additionally,
independent whole-brain analyses from structural
neuroimaging studies related the gray-matter volume GMV of the vmPFC,
more consistently than any other
brain region, to indices of
social competence and social network complexity,
among the most complicated decision that humans and monkeys take
\citep{behrens2009computation}.
%

@@

RL is a problem-solving technique in which, through their interactions with an external environment,
an agent learns to reach goals, optimize reward signals, or minimize costs
in an iterative trial-by-error fashion.
At any moment, each taken \textit{action} $a$ triggers a changes in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by a real-valued \textit{reward}
$r = r(s, a)$ (or regret!, since it can be negative) collected by the
agent.
In this view, the environment is partially controlled by
the action of the agent and the reward should be thought
of as satisfaction -- or punishment -- accompanying the execution of
an action.
Action may be delayed to achieve substantial increases in expected reward
that can increase over time.
The environment is generally taken as stochastic,
that is, unpredictable, changing in random ways. In addition,
it is only \textit{partially observed},
in the sense that only part of the current state is observed by
the agent \citep{starkweather2017dopamine}.
This unpredictability of the environment
is realistic in a model which sets out
to explain DMN function. Indeed, it is known that the more the external world is predictable,
the more mental activity becomes detached from the actual sensory inputs
\citep{antrobus1966studies, pope1978regulation},
the more stimulus-independent thoughts occur, and
the higher the neural activity in the DMN.
Conversely, the more the currently executed task is unknown and unpracticed,
the lower the DMN activity
\citep{filler1973daydreaming, teasdale1995stimulus}.
%
These ``offline'' processes may however contribute to
optimizing control of the organism.
Such a framework of DMN funtion based on reinforcement learning
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules
\citep{dayan2008decision}.
The implication of the DMN in a diversity of
particularly advanced cognitive processes
can be parsimoneously explained as prediction error minimization
coupled with probabilistic simulations in form of mental scenes,
thus calibrating action portfolios and
maximizing reward outcome at various time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.

\subsection{Partially Observable Markov Decision Processes}
It is our view that at a sufficiently coarse scale, the brain -- and therefore the DMN in particular --
is a physical system governed by the laws of
physics and can be locally described by Markov processes. Indeed, as noted in \cite{tegmark2016improved},
any system obeying the laws of classical physics can be accurately modeled as a Markov process as long as the time
step is sufficiently short, defining $s(t)$, the state at time $t$, as the position in phase space.
If the process has $memory$ such that the next state depends not only on the current state but also on some finite
number of past states.
Rational probabilistic planning can be reformulated
as a standard memoryless Markov process by simply expanding the
definition of the state $s$ to include elements of the past.

In artificial intelligence (AI), a popular computational model for
multi-step decision processes in such an environment are so-called
\textit{Partially Observable Markov Decision Processes (POMDPs)}.
A POMDP formalizes a decision process in which it is assumed that environment dynamics are determined by a Markov process,
but the agent cannot directly observe the underlying state.
Instead, the agent tries to optimize a \textit{subjective} reward
signal (i.e., is likely to be different for another agent in the same state).
This is performed alongside each transition in the environments state-space, and maintains a probability distribution over the set of possible
states, based on a set of observations and observation probabilities. This is a minimal amount of assumptions that can be made about an environment,
and is characteristic of so-called \textit{model-free RL}, the approach adapted by our work.

@@

In contrast, \textit{model-based RL} would start off with some mechanistic assumptions about the dynamics of world.
These can be assumptions about the physical laws governing the agent's world or constraints on the state space and transitions between states.
For instance, if a rat in a maze knows
that standing still will produce no change in the environment, and in particular will not eventually lead to finding the food.
% A core property of human intelligence has been proposed underly to improve
% on an expected utility outcomes as a strategy for action choice in uncertain
% environments \citep{gershman2015computational}.

In our present model-free RL framework,
a rat might represent such knowledge about the world as follows:
\begin{itemize}
\item $r(s, \text{``stand still''}) = 0$ if $s$ does not correspond to a cell / chamber containing food.
\item $p(s'|s,\text{``stand still''}) = 1$ if $s'=s$ and $0$ otherwise.
\item etc.
\end{itemize}

@@

\subsection{Long-term rewards and action policies}
The behavior of the agent is governed by some kind of \textit{policy},
which maps states of the world to
a set of candidate actions to perform given this state. Starting a time $t=0$,
a policy $\pi$ generates a trajectory of action cascades as follows
\begin{eqnarray*}
  \begin{split}
    \textbf{ observe transition: }s_1 &\sim p(s|s_0,a_0)\textbf{ and collect rewards }R_0 = r(s_0, a_0)\\
    \textbf{choose action: }a_1 &\sim \pi(a|s_1)\\
    \textbf{ observe transition: }s_2 &\sim p(s|s_1,a_1), \textbf{ and collect rewards }R_1 = r(s_1, a_1)\\
    \ldots\\
    \textbf{choose action: }a_{t} &\sim \pi(a|s_{t})\\
    \textbf{ observe transition: }s_{t+1} &\sim p(s|s_{t},a_{t}), \textbf{ and collect rewards }R_{t} = r(s_{t}, a_{t})\\
    \ldots
  \end{split}
\end{eqnarray*}
Since an action taken in the present moment $t$
might have repercussions long into the future, it turns out that the
quantity to optimize is not the instantaneous rewards $r(s, a)$, but a
cumulative reward estimate which takes into account the future. A standard for
modeling this accumulation is the time-discounted \textit{cumulative reward}% function
\begin{equation}
  \label{eq:cumr}
  G^\pi = \sum_{t=0}^{\infty}\gamma^{t}R_t = R_0 + \gamma R_1 + \gamma^2 R_2 + \ldots + \gamma^tR_t + \ldots
\end{equation}
This random variable\footnote{Random as it depends both on the environment's dynamic and the
  policy $\pi$ being played (which can be stochastic).}  measures the cumulative reward of
following a policy $\pi$.
\begin{mdframed}
  \textbf{Where is ``value'' buffered in the brain ? You mean where in the DMN!
  then the answer is in the vmPFC which is connected to the NAc}
  The amygdala is known to be involved in significance evaluation) and
  the nucleus accumbens (NAc) is involved in reward evaluation.
\end{mdframed}
  The  goal of RL is then to adapt this policy so as to  maximize $G^\pi$. In
  the definition of cumulative reward $G^\pi$ above, the constant $\gamma$ $(0 \le \gamma < 1$) is the reward \textit{discount factor}.
Setting $\gamma = 0$ corresponds to a perfectly myopic agent who is solely concerned about its
immediate rewards.
Such an agent would have no horizon for the future,
which is not compatible with long-term goal planning
as potentially implemented in the DMN.
To allow a learning process to arise,
it is necessary that $0 < \gamma < 1$.
% \footnote{In which case we can take $T \rightarrow +\infty$, provided the reward function $r$ is bounded.}.
Such an agent is thus better emphasizing consideration of rewards expected in
a short future and paying less attention to
rewards expected at a later time.
% given that there is more uncertainty in the farsighted future.
The reward discount factor $\gamma$ can then be seen as calibrating
the delay discounting behavior of the intelligent agent,
that is, the decision related to reward-delay trade-off schedules.
It is also important to appreciate that a stochastic policy estimation
is more advantageous in many RL settings.
For instance, a deterministic strategy in playing rock-paper-scissors can
be easily exploited and a uniformly random choice is optimal.

\subsection{The Q-value: A goodness measure for action choices}
The purpose of the DMN can then be
formulated as finding a policy $\pi$ for the agent which maximizes the
expected cumulative value of an state-action pair $(a,s)$, also known as the \textit{Q-value}, namely
\begin{equation}
  \label{eq:qval}
  Q^{\pi}(s,a) = \mathbb E [G^\pi|s_0=s,a_0=a,a_1 \sim \pi,a_2 \sim \pi,\ldots].
\end{equation}
In other words,
the Q-value $Q^{\pi}(s,a)$ corresponds to the expected reward
over all possible action trajectories, in which
the agent sets out in the environment in state
$s$, chooses action $a$,
and then follows the policy $\pi$ thereafter to select future actions.
%% Given that each $\pi$ correspondence to an exhaustive description of
%% behavioral patterns, $Q^{\pi}(s_t,a_t)$ corresponds to the epistemic value of
%% the set of actions given states in finite time.
For the brain,
$Q^{\pi}(s, a)$ defined in \eqref{eq:qval} provides the subjective
utility of executing a specific action; it answers the question
"What is the expected utility of taking action $a$ in this situation ?".
When a choice of actions is available, the $Q^{\pi}(s,a)$
offers a formalization of optimal behavior,
potentially driven by the DMN.

\subsection{Computing optimal behavior}
By construction, maximizing the expectation in  \eqref{eq:qval} is computationally intractable in general.
This is due to the enormous number of possible action
trajectories the agent might need to
consider simultaneously.
% \begin{mdframed}
%   \textbf{N.B.:}
%   If the Bellman-based computational ideas have a corresponding
%   neurobiological implementation, this provides a framework of how
%   optimal behavior could be computable in the human brain.
% \end{mdframed}
One popular solution concept is called \text{Q-learning}. It remedies
the issue by only considering the sub-family of
\textit{deterministic policies}, which map each state to the single best action to
take at the state. Such policies define deterministic functions from states to actions.

\subsubsection{Q-learning} Q-learning (Watkins \& Dayan, 1992), generates a
\textit{greedy policy}
\begin{equation}
  \pi(s) = \argmax_{a \in \mathcal A}{Q}(s, a).
  \label{eq:qlearning}
\end{equation}
% where $\tilde{Q}(s, a)$ is the current estimate of the state-action value
% table.
The \textit{Bellman equation} \citep{sutton1998reinforcement} then takes the simple form
\begin{equation}
  \begin{split}
    Q^\pi(s, a) &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma \max_{a' \in \mathcal A}Q^\pi(s', a')]
    \\
    &= \mathbb E_{s' \sim p(s'|s,a)} [r(s,a) + \gamma Q^\pi(s', \pi(s'))].
  \end{split}
  \label{eq:bellman}
\end{equation}
This \textit{fixed-point} equation provides a recursive decomposition of optimal behavior.
A complicated dynamic programming optimization can be
broken into simpler sub-problems at different time points,
which has often been closely related to the medial prefrontal part of the DMN
\citep{koechlin1999role, braver2002role}.
Using the Bellman equation, each state can be associated with a certain value
to guide action towards a better state, thus improving on the agent's current policy.
% It is a fixed-point equation for the deterministic policy $\pi$.
Notice that in \eqref{eq:bellman}, the sampling is done only over things which
depend on the environment, and so can be learned off-policy by observing state transitions
triggered by another behavioral policy, which can be stochastic.

\subsubsection{Efficient learning via value approximation and experience replay}
%As already mentioned in the previous section, Q-learning optimizes over the class of
% deterministic policies of the form \eqref{eq:qlearning}.
A successful RL theory needs ways to approximate the state-action value function
to allow for scaling to large state and action spaces.
In state $s \in \mathcal S$, the agent takes an action $a \in \mathcal A$
% (for example, $a$ = ``Read'', or ``Run'', or ``Laugh'', or ``Sympathize'', or ``Empathize'', etc.)
by sampling from its current policy matrix, and collects a reward $r$,
% (measured on their internal currency of worldly values ?),
and the environment transitions to a new state $s' \in \mathcal S$.
\textbf{XXX Please try to smoothen the previous sentence}
At the end of
this step, a new \textit{experience} $e = (s,a,r,s')$ is produced; this represents an exemplar
behavior of the agent and is recorded in replay memory buffer:
$D \leftarrow D \cup \{e\}$ (possibly discarding the oldest entries to make space).
Now, at iteration $k+1$, replay consists in sampling (uniform or importance-weighted
\footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.} ?)  mini-batches of experiences
$(s, a, r, s') \sim \mathcal U(D)$ from the replay memory $\mathcal D$
and trying to approximate
the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the Bellman equation \eqref{eq:bellman}, namely
$r + \gamma \max_{a'} Q(s', a'|\theta^Q_k)$, with the output of a parametrized regression model $(s,a,r,s')
\mapsto {Q}(s, a|\theta^Q_{k})$.
Here, $Q(.,.|\theta^Q_{k})$
\textbf{XXX just double-checking, are the dots int he Q() correct?}
is an approximator for the $Q$-value operator,  parameterized by weights $\theta^Q_{k+1}$
(synaptic strengths, etc.).
